Hydrous melting and partitioning in and above the mantle transition zone: insights from water-rich MgO-SiO$_2$-H$_2$O experiments

### 2016-05-10 initial_submission/MSH.tex
Updated to revised version

\journal{Geochemica and Cosmochemica Acta}

@@

The equilibrium constant is related to the energy $\Delta G_{(1)}$ required to protonate two bridging oxygens with a water molecule
\begin{equation}
K_{(1)} = \exp\left(\frac{-\Delta G_{(1)}}{RT}\right)
\end{equation}

@@

These models can be used to calculate melt compositions by equating the gibbs free energy of the solid $G^s_{solid}$ with the chemical potential of the component with the composition of that solid in the melt $\mu^s_{melt}$:
\begin{equation}
G^s_{solid} = \mu^s_{melt} = G^s_{melt} - RT \ln a^s_{melt}
\label{eqn:equilibrium}
\end{equation}
In this study, we use the thermodynamic dataset of \cite{SLB2011} to calculate $G^s_{solid}$ for periclase, forsterite and its high pressure polymorphs, orthoenstatite, high pressure clinoenstatite, stishovite and coesite. A modified version of the molecular dynamics-derived anhydrous melt model of \cite{DKS2013} is used to calculate $G^s_{melt}$, and the \cite{SS1985} model is used to calculate $a^s_{melt}$. The calculations in the work were undertaken with the mineral physics toolkit Burnman \citep[available from \url{https://geodynamics.org/cig/software/burnman/}][]{CHRU2014}.

@@

\begin{figure}[ht!]
  \centering
      %\includegraphics[width=0.8\textwidth]{figures/forsterite}
      \caption{The forsterite-water phase diagram at 13 GPa, based on current experimental results. The thermodynamics of pure forsterite are taken from the literature \citep{SLB2011}, and the properties of the anhydrous melt are from the mixing model described in the text. At high temperatures, the forsterite liquidus is metastable with respect to periclase + melt.}
  \label{fig:foH}
\end{figure}
\clearpage

@@

\begin{figure}[ht!]
  \centering
      %\includegraphics[width=0.8\textwidth]{figures/enstatite}
  \caption{The clinoenstatite-water phase diagram at 13 GPa, based on current experimental results. The thermodynamics of pure high pressure clinoenstatite are taken from the literature \citep{SLB2011}, and the properties of the anhydrous melt are from the mixing model described in the text.}
  \label{fig:eoH}
\end{figure}

>>>>

\journal{Geochimica et Cosmochimica Acta}

@@

The equilibrium constant is related to the energy $\Delta G_{(1)}$ required to create two terminal OH groups from a bridging oxygen and a water molecule
\begin{equation}
K_{(1)} = \exp\left(\frac{-\Delta G_{(1)}}{RT}\right)
\end{equation}

@@

These models can be used to calculate solid-melt equilibria by equating the Gibbs free energy of the solid $G^s_{solid}$ with the chemical potential of the melt component with the same composition $\mu^s_{melt}$:
\begin{equation}
G^s_{solid} = \mu^s_{melt} = G^s_{melt} - RT \ln a^s_{melt}
\label{eqn:equilibrium}
\end{equation}
In this study, we use the thermodynamic dataset of \cite{SLB2011} to calculate $G^s_{solid}$ for periclase, forsterite and its high pressure polymorphs, orthoenstatite, high pressure clinoenstatite, coesite and stishovite. A modified version of the molecular dynamics-derived anhydrous melt model of \cite{DKS2013} is used to calculate $G^s_{melt}$, and the \cite{SS1985} model is used to calculate $a^s_{melt}$. The calculations in the work were undertaken with the mineral physics toolkit Burnman \citep[available from \url{https://geodynamics.org/cig/software/burnman/}][]{CHRU2014}.

@@

\begin{figure}[ht!]
  \centering
      %\includegraphics[width=0.8\textwidth]{figures/forsterite}
      \caption{The forsterite-water phase diagram at 13 GPa, based on current experimental results. The thermodynamics of pure forsterite are taken from the literature \citep{SLB2011}, and the properties of the anhydrous melt are from the mixing model described in the text. At high temperatures, the forsterite liquidus is metastable with respect to periclase + melt. High temperature data come from the study of \cite{NDMF2015}.}
  \label{fig:foH}
\end{figure}
\clearpage

@@

\begin{figure}[ht!]
  \centering
      %\includegraphics[width=0.8\textwidth]{figures/enstatite}
  \caption{The clinoenstatite-water phase diagram at 13 GPa, based on current experimental results. The thermodynamics of pure high pressure clinoenstatite are taken from the literature \citep{SLB2011}, and the properties of the anhydrous melt are from the mixing model described in the text. High temperature data come from the study of \cite{NDMF2015}.}
  \label{fig:eoH}
\end{figure}
